First steps towards a formalization of Forcing

### 2018-06-25 LSFA/intro.tex
volvemos por laburo de Manu
Modificación mínima

Assume that $M$ is a ctm of $\ZFC$. It is instructive to sketch
G\"odel's argument of the relative consistency of $\CH$: he showed
that $M$ contains a minimal submodel $L^M$ of the same ``height''
(i.e.\ having the same ordinals) that satisfies $\ZFC+\CH$. The sets
in $L^M$ are called \emph{constructible} and are in a sense
``definable.'' In fact, there is a first-order formula $L$ such that
$L^M = \{x\in M : M\models L(x)\}$. To show that
$L^M\models \ZFC+\CH$, one uses that the $\ZFC+\CH$ holds in $M$.

@@

A central part of this formalization project involves showing that
$\ZFC$ holds in the generic extension. This is most relevant since
forcing is essentially the only known way to \emph{extend} models of
$\ZFC$ (while preserving ordinals). The most difficult step to achieve
this goal is to define the \emph{forcing relation}, that allows to
connect satisfaction in $M$ to that of $M[G]$; this is needed to show
that the Powerset axiom and the axiom schemes of Separation and
Replacement hold in $M[G]$. In
Section~\ref{sec:pairing-generic-extension}
 we tackle the Pairing
Axiom. This does not require the forcing relation, but provides an
illustration of the use of names. The development can be downloaded
from \url{https://cs.famaf.unc.edu.ar/~pedro/forcing/}.

>>>>

It is instructive to sketch G\"odel's argument of the relative
consistency of $\CH$: Assume that $M$ is a ctm of $\ZFC$, G\"odel
showed that $M$ contains a minimal submodel $L^M$ of the same
``height'' (i.e.\ having the same ordinals) that satisfies
$\ZFC+\CH$. The sets in $L^M$ are called \emph{constructible} and are
in a sense ``definable.'' In fact, there is a first-order formula $L$
such that $L^M = \{x\in M : M\models L(x)\}$. To show that
$L^M\models \ZFC+\CH$, one uses that the $\ZFC+\CH$ holds in $M$.

@@

A central part of this formalization project involves showing that
$\ZFC$ holds in the generic extension. This is most relevant since
forcing is essentially the only known way to \emph{extend} models of
$\ZFC$ (while preserving ordinals). The most difficult step to achieve
this goal is to define the \emph{forcing relation}, that allows to
connect satisfaction in $M$ to that of $M[G]$; this is needed to show
that the Powerset axiom and the axiom schemes of Separation and
Replacement hold in $M[G]$. In
Section~\ref{sec:pairing-generic-extension}
 we tackle the Pairing
Axiom. This does not require the forcing relation, but provides an
illustration of the use of names. The development can be downloaded
from \url{https://cs.famaf.unc.edu.ar/~mpagano/forcing/}.
